О множествах точек на плоскости с целочисленными расстояниями

### 2019-05-30 pps-linear-bound/pps-linear-lower-bound.tex
Typo fix
Just labels
The main construction of the bound

Basing on the exact values of $d(2,n)$ for $ 3 \geq n\leq 122$ and the estimate $d(2,123)>10000$
\cite{kurz2008bounds}, we derive
\begin{observation}
	The inequality
	\begin{equation}
		d(2,n) \geq 3^{1/4}\cdot2^{-3/2} \cdot n
	\end{equation}
	holds for $4 \leq n \leq 21491$.
\end{observation}

Performing some trivial manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
one can make the following
\begin{observation}
	For $n \geq 21492$ we have
	\begin{equation}
		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{k-2}}
		,
	\end{equation}
	where
	\begin{equation}
		\beta = \frac{1}{\sqrt{21490}} + \sqrt{ \frac{2}{\sqrt{3}} + \frac{1}{21490} } < 1.07464
		.
	\end{equation}
\end{observation}

We can improve the constant in Theorem~(TODO: ref) if new exact values of $d(2,n)$ are found;
however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$.
To be precise, we have the following

>>>>

Basing on the exact values of $d(2,n)$ for $ 3 \leq n\leq 122$ and the estimate $d(2,123)>10000$
\cite{kurz2008bounds}, we derive
\begin{observation}
	\label{obs:4_leq_n_leq_21491}
	The inequality
	\begin{equation}
		d(2,n) \geq 3^{1/4}\cdot2^{-3/2} \cdot n
	\end{equation}
	holds for $4 \leq n \leq 21491$.
\end{observation}

Performing some trivial manipulations with the upper bound in Theorem~\ref{thm:varphi_k_bounds},
one can make the following
\begin{observation}
	\label{varphi_n_where_n_geq_21492}
	For $n \geq 21492$ we have
	\begin{equation}
		\varphi_n \leq \varphi_{n-1} \leq \frac{\beta}{\sqrt{n-2}}
		,
	\end{equation}
	where
	\begin{equation}
		\beta = \frac{1}{\sqrt{21490}} + \sqrt{ \frac{2}{\sqrt{3}} + \frac{1}{21490} } < 1.07464
		.
	\end{equation}
\end{observation}

In [TODO:cite], TODO: theorem about sublinear quantity of points on the line.
However, we need the values of all the constants in such an estimate.

\begin{definition}
	A $\rho(k,M_1 M_2)$-curve, where $-|M_1 - M_2| < k < |M_1 - M_2|$ and $k$ is integer, is the set of points $N$,
	such that for given points $M_1$ and $M_2$
	the equality $|N - M_1| - |N-M_2|$ holds.
\end{definition}
So, $\rho(k,M_1 M_2)$-curve is a branch of a hyperbola for $k\neq 0$ and the bisector of line segment $M_1 M_2$ for $k=0$.

\begin{observation}
	\label{obs:rho_curves}
	If points $M_1,M_2,M_3,M_4$ are situated on a straight line,
	and equality $|M_1 - M_2| = |M_3 - M_4|$ holds but line segments $M_1 M_2$ and $M_3 M_4$ do not coincide,
	then for a fixed $k$ the $\rho(k,M_1 M_2)$-curve and $\rho(k,M_3 M_4)$-curve
	do not intersect.
\end{observation}

\begin{lemma}
	\label{lem:2k-1_segments}
	Let $M \in \mathfrak{M}(2,n)$ for some $n$ and let $m$ be a straight line.
	Then for every $k\in\mathbb{N}$ there are at most $2k-1$ segments $\Delta_i \subset m$,
	such that the length of $\Delta_i$ is exactly $k$.
\end{lemma}
\begin{proof}
	Consider a point $N\in M \setminus m$.
	Then for each $\Delta_i$ there is a $\rho(n_i,\Delta_i)$-curve containing $N$.
	Due to Observation~\ref{obs:rho_curves}, all $n_i$ are distinct;
	otherwise $\rho(n_i,\Delta_i)$-curve and $\rho(n_i,\Delta_j)$-curve, $j\neq i$, do not intersect.
	There can be only $2k-1$ different values for $n_i$,
	so there are at most $2k-1$ distinct segments $\Delta_i$.
\end{proof}

TODO: different vs. distinct

\begin{lemma}
	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n$.
	Let $|\Delta \cap M| = n^2 + 1$.
	Then $l \geq \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n$
	.
\end{lemma}

TODO: $|\cdot|$ for length, $\#$ for cardinality?

\begin{proof}
	$n^2+1$ points, including the ends of $\Delta$, split the segment $\Delta$ into $n^2$
	sequential segments $\Delta_i$.
	Due to Lemma~\ref{lem:2k-1_segments}, there is at most one segment of length 1,
	at most three segments of length 2 etc.
	The following two values of sums conclude the proof:
	\begin{equation}
		1 + \sum_{k=1}^n (2k-1) = n^2 + 1
		,
	\end{equation}
	\begin{equation}
		\sum_{k=1}^n k(2k-1) = \frac{2}{3}n^3+\frac{1}{2}n^2-\frac{1}{6}n
		.
	\end{equation}
\end{proof}

Now we will estimate the length of a line segment, which shares an arbitrary number of points
with an integral point set.

\begin{lemma}
	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n>122$.
	Let $|\Delta \cap M| = k > 10000$.
	Then $l\geq \gamma_1 k^{3/2}$,
	where

	TODO!
\end{lemma}

This lemma leads to the following observation.

\begin{observation}
	\label{obs:estimate_points_on_straight_line}
	Let $\Delta$ be a straight line segment of length $l$, $M \in \mathfrak{M}(2,n)$ for some $n>21491$.
	Let $|\Delta \cap M| = k > 10000$.
	Then $k \leq \gamma_2 b + 6$,
	where

	TODO!
\end{observation}

Now we are ready to proof the main theorem of the section.

\begin{theorem}
	\label{thm:main_estimate}
	If $n\geq 4$, then $d(2,n) \geq \gamma (n - 2)$,
	where
	\begin{equation}
		\gamma = ... < \frac{3^{1/4}}{2^{3/2}}
	\end{equation}
	TODO!
\end{theorem}

\begin{proof}
	For $4 \leq n \leq 21491$, the claim of the theorem follows immediately from Observation~\ref{obs:4_leq_n_leq_21491}.
	Let us consider $M\in \mathfrak{M}(2,n)$, $n \geq 21492$, $\operatorname{diam} M = b$.
	Lemma~\ref{lem:square_container} yields that $M$ is situated is a square with side length $b$.
	Let $M_1, M_2, M_3, M_4$ be points of $M$ such that the distances $|M_1-M_2|$ and $|M_3-M_4|$
	are minimal in $M$ ($M_2$ and $M_3$ may be the same point).
	Then $|M_1-M_2| \leq b\varphi_{n-1}$ and $|M_3-M_4| \leq b\varphi_{n-1}$.
	Due to Lemma~\ref{lemma:quadr_diag_edges}, open line segments $M_1 M_2$ and $M_3 M_4$ do not intersect
	(otherwise they are not minimal).

	Let $C = cr(M_1 M_2) \cap cr(M_3 M_4)$.
	Each point of $n\in M$ satisfies one of the following conditions:

	a) $N$ belongs to $C$~--- totally at most $\gamma_2 b + 6$ points due to Lemma~\ref{lem:intervals_cross} and Observation~\ref{obs:estimate_points_on_straight_line};

	b) $N$ belongs to the intersection of one of $|M_1 - M_2| - 1$ hyperbolas
	with one of $|M_3 - M_4| - 1$ hyperbolas~--- totally at most $4 (|M_1 - M_2| - 1)(|M_3 - M_4| - 1)$ points;

	c) $N$ belongs to the intersection of one of $|M_1 - M_2| - 1$ hyperbolas and $cr(M_3 M_4)$~---
	totally at most $4 (|M_1 - M_2| - 1)$ points;

	d) $N$ belongs to the intersection of one of $|M_3 - M_4| - 1$ hyperbolas and $cr(M_1 M_2)$~---
	totally at most $4 (|M_3 - M_4| - 1)$ points;

	(see~\cite[]{erdos1945integral} for details).

	Summing up the cases above, we get the following estimate:
	\begin{equation}
		\label{eq:estimate_n_varphi_n_squared}
		n \leq 4 b^2 \varphi_{n-1}^2 - 4 + \gamma_2 b + 6
		.
	\end{equation}
	Observation~\ref{varphi_n_where_n_geq_21492} turns estimate~\eqref{eq:estimate_n_varphi_n_squared}
	into the following:
	\begin{equation}
		\label{eq:estimate_n_sqrt_n-2}
		n-2 \leq 4 b^2 \frac{\beta^2}{n-2} + \gamma_2 b
		,
	\end{equation}
	which obviously leads to inequality
	\begin{equation}
		1 \leq 4\beta^2 \left(\frac{b}{n-2}\right)^2  + \gamma_2 \frac{b}{n-2}
		.
	\end{equation}
	Let us denote $\lambda = b/(n-2)$ and solve the following square inequality with respect to $\lambda$:
	\begin{equation}
		\label{eq:square_inequality_lambda}
		4\beta^2 \lambda^2  + \gamma_2 \lambda - 1 \geq 0
		.
	\end{equation}
	The discriminant is $\gamma_2^2 + 16 \cdot \beta^2$,
	so we obtain the following estimate:
	\begin{equation}
		\frac{b}{n-2} = \lambda \geq \frac{-\gamma_2 + \sqrt{\gamma_2^2 + 16 \cdot \beta^2}}{8\beta^2}
		.
	\end{equation}
	Calculating the expression above for our $\beta$ and $\gamma_2$, we conclude the proof.
\end{proof}

We can improve the constant in Theorem~\ref{thm:main_estimate} if new exact values of $d(2,n)$ are found;
however, the obtained constant is bounded by $3^{1/4}\cdot2^{-3/2}$.
To be precise, we have the following
